Seagrass mapping and monitoring along the coasts of Crete, Greece

### 2011-01-18 MScDiplom_Lemenkova.tex
added flowchart, p.25

\textbf{Software and utilities used in the current work}:\\
\begin{itemize}
	\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
	\item \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
	\item WASI Water Color Simulator, for modelling optical water properties under changing environment
	\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and GPS for data capture during the fieldwork
	\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
	\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
	\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
	\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
	\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data preprocessing and preliminary statistical evalution
	\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
	\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
	\item \href{http://staruml.sourceforge.net/en/}{StarUML}: modelling flowchart diagrams for work algorithms
\end{itemize}
\pagebreak

@@

\subsection{Review of the collected data}
The remote sensing imagery were collected for the Crete island area covering the research period and
area, enabling the field data to be used for the calibration and validation. The available aerial and
satellite imagery are commonly used for the mapping of seagrass landscapes and their application is
proven by various research papers. The imagery include satellite multi-spectral imagery (Landsat-TM,
ETM+) and aerial imagery from the Google Earth.
The overview of the collected data enables to summarize their following types:
\begin{itemize}
	\item[] 1. Optical spectra of P.oceanica, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig. \ref{fig:23})
	\item[]2. Aerial imagery from the \href{http://www.google.com/earth/index.html}{Google Earth}
	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:68})
	\item[]4. Results of underwater videometric measurements of the Olympus cameras made during the ship route
\end{itemize}
The available satellite imagery (Fig.\ref{fig:68}) and aerial images are read into the ArcGIS project.
The available broadband and hyperspectral remote sensing data are used for the mapping of the
seagrass in shelf areas no deeper than 4.0 meters, for the environmental monitoring in order to detect
the dynamics in changes of the seagrass distribution along Crete during the past 10 years using
different satellite images for 2000-2010.
The spatial resolution of Landsat ETM+ image is 30 m in the visible and near infrared bands (bands
1-5 and 7); the spatial resolution of ASTER 15 m for the visible and near-infrared bands. IKONOS
acquires data in 3 visible channels and NIR, with spatial resolution of 1-4 meters. ASTER and
IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
research includes fieldwork in-situ measurements of the seagrass distribution along the northern and
south-eastern coasts of Crete in chosen locations.\\
The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
the seafloor have been received by the means of the optical sensors Trios-RAMSES Hyperspectral
UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig. \ref{fig:10}), both adjusted for the measurements of the irradiance and radiance (see Appendices: Tab. \ref{tab:18}, Tab. \ref{tab:19}).
The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:49}) and 430-650 for the water radiance (Fig.\ref{fig:47}). 
Afterwards, the measured values of the
radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
properties of the sea water and bottom cover types (Fig.\ref{fig:23}). The spectral range of radiance
cover diapason of 320-950 nm, and ittadiance measurements are covered in the interval of 280-500
nm, which is suitable for a characteristics of seagrass reflectance.
Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:43}, or \ref{fig:45}, enlarged) represent several series of
the measurements. The values of the spectral reflectance are received from the
computations of these values using Excel mathematical formulae. The graphs shown on Fig.\ref{fig:47}, Fig. \ref{fig:48}, Fig. \ref{fig:49} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively. Graph on Fig \ref{fig:44} displays statistuical analysis of the measured sets of observations, i.e. shows the midspread of the statistical quartiles (Q1-Q3), mean and extreme values.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{GNU-17.jpg}
	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete}
	\label{fig:23}
\end{figure}
\pagebreak

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-18.jpg}
	\caption{Multiplot graph showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete. Gnuplot. \\Two complimentray graphs below show the results of the statistical analysis}
	\label{fig:26}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-21.jpg}
	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete. Gnuplot. \\Two complimentray graphs below show the results of the statistical analysis}
	\label{fig:27}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-20.jpg}
	\caption{Statitical comparison of spectral reflectances of \textit{P.oceanica} and carbonate sand. Selected measurements sets}
	\label{fig:28}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.40]{GNU-19.jpg}
	\caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and 			measured values (dotted): multiplot of measurement sets 200-300}
	\label{fig:33}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.45]{GNU-13.jpg}
		\caption{Multiplot of spectral reflectance of P.oceanica. Series 1-100. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
			mean value within the range (red bold dots). Gnuplot visualization­}
		\label{fig:35}
	\end{center}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[scale=0.5]{GNU-16.jpg}
		\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
			mean value within the range (red bold dots). Gnuplot visualization­}
		\label{fig:36}
	\end{center}
\end{figure}

@@

\begin{figure}
	\centering
	\subfloat [small-scaled (1:54,000)]{\label{fig:35a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
	\subfloat [enlarged (1:10,000)]{\label{fig:35b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
	\caption{Seafloor types of the Arina beach: \textit{P.oceanica} sand, rocks}
	\label{fig:37}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.40]{Fig-36.jpg}
		\caption{Results of the image classification in Erdas Imagine: seagrass distribution in Bali area, Crete}
		\label{fig:38}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.25]{Fig-37.jpg}
		\caption{Google aerial images incorporated into the GIS project: fragment of ArcMap layout}
		\label{fig:39}
	\end{center}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfloat [Meadows of the seagrass. Source: Googel Earth]{\label{fig:38a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
	\subfloat [Patches (mattes) of the seagrass. Source: in-situ videometric measurements]{\label{fig:38b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
	\caption{Variations in spatial structure of the seagrass landscapes}
	\label{fig:40}
\end{figure}

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.25]{Fig-39.jpg}
		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the ArcGIS project}
		\label{fig:41}
	\end{center}
\end{figure}

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.20]{App-1.jpg}
		\caption{Values of spectral reflectances of seagrass P.oceanica (fragment)}
		\label{fig:42}
	\end{center}
\end{figure}
\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.20]{GNU-401-420.jpg}
		\caption{Spectral reflectance of P.oceanica. Measurement series 401-420}
		\label{fig:43}
	\end{center}
\end{figure}
\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.20]{GNU-401-420-candles.jpg}
		\caption{Statistical analysis of the measurement data: spectral reflectance of the P.oceanica. 
		Visualisation of the interquartile ranging. Example of data set 401-420. Shown midspread of  statistical quartiles Q1 and Q3, min and max values within the range.}
		\label{fig:44}
	\end{center}
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.20]{GNU-enlarged-M-401-420.jpg}
		\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
		Visualisation of the interquartile ranging and plotted together with measurement data.}
		\label{fig:45}
	\end{center}
\end{figure}
\pagebreak
%%%

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral 			reflectance of \textbf{P.oceanica}. SPSS}
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline\hline
			\textbf{R} & \textbf{R square} & \textbf{Adjusted R square} & \textbf{Std. Error of the Estimate} \\ \hline\hline
			.277 1 & .077 & .068126692 & .462 \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:5}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{P.oceanica}. SPSS}
	\begin{center}
		\begin{tabular}{|c|c|c|c| c|c|}
			\hline\hline
			& \textbf{Sum of squares} & \textbf{df} & \textbf{Mean Square} & F & Sig.\\ \hline\hline
			Regression & 1.810 & 1 & 1.810 & 8.477 & .004\\ \hline
			Residual & 21.781 & 102 & .214 & & \\ \hline
			Total & 23.592 & 103 & & & \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:6}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. SPSS}
	\begin{center}
		\begin{tabular}{| c | c | p{2cm}| c | c | c |}
			\hline\hline
			& \multicolumn{2}{|c|}{Unstandardized Coefficients}\\
			\cline{2-3}
			& B & Std.Error & Stand. Coef. Beta & t & Sig.\\ \hline\hline
			wavelength & -.001 & .000 & -.277 & -2.912 & .004 \\ \hline
			constant & .464 & .121 & & 3.821 & .000 \\ \hline
		\end{tabular}
	\end{center}
	The underlying process assumed is independence (white noise).\\
	 Based on the asymptotic chi-square approximation.
	\label{tab:7}
\end{table}
\pagebreak

\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=0.20]{GNU-Stat-M-401-420.jpg}
	\caption{Fragment of the statistical analysis of the P.oceanica reflectance. Example of data set 401-420. 
		Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means.}
	\label{fig:46}
	\end{center}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.25]{GNU-Radiance_Bezier.jpg}
	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation}
	\label{fig:47}
\end{figure}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline\hline
		\textbf{Source of Variation} & \textbf{SS} & \textbf{df} & \textbf{MS} & \textbf{F} & \textbf{P value} & \textbf{F crit} \\ \hline
		Between Groups & 373841.7048 & 2 & 186920.8524 & 407.85359 & 1.11677 & 3.01153 \\ \hline
		Within Groups & 261233.1668 & 570 & 458.3038014 & & & \\ \hline
		Total &  635074.8716 & 572 & & & & \\ \hline
	\end{tabular}
	\end{center} 
	P more than .05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
	\label{tab:8}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|}
		\hline\hline
		\textbf{Groups} & \textbf{Count} & \textbf{Sum} & \textbf{Average} & \textbf{Variance} \\ \hline\hline
		Column 1 & 191 &	760.126692 & 3.9797209 & 18.6948046 \\ \hline
		Column 2 & 191 &	4527.214218 & 23.7026922  & 251.4863978 \\ \hline
		Column 3 & 191 &	12465.19972 & 65.2628257 & 1104.7302015 \\ \hline
	\end{tabular}
	\end{center}
	P$>$0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
 	\label{tab:9}
\end{table}
\pagebreak

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-Irr-Splines.jpg}
		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in GNUplot­}
		\label{fig:48}
	\end{center}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.25]{GNU-Irr.jpg}
	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in GNUplot}
	\label{fig:49}
\end{figure}
\pagebreak
\newpage
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-gamma-rad.jpg}
		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
		\label{fig:50}
	\end{center}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{G-normal-dist.jpg}
	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
	\label{fig:51}
\end{figure}

%%% Gretle start
\begin{table}
	\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16. Gretl}
	\begin{center}
	Model 2: Logistic, using observations 1:01--8:23 ($T$ = 191)\\
	Dependent variable: v16\\
	$\hat{y} = 1 / (1 + e^{-X\hat{\beta}})$

>>>>

\textbf{Software and utilities used in the current work}:\\
\begin{itemize}
	\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
	\item \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
	\item \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}, for modelling optical water properties under changing environment
	\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and GPS for data capture during the fieldwork
	\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
	\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
	\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
	\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
	\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data preprocessing and preliminary statistical evalution
	\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
	\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia}: modelling flowchart diagrams for work algorithms
\end{itemize}
\pagebreak

@@

\subsection{Review of the collected data}
The remote sensing imagery were collected for the Crete island area covering the research period and
area, enabling the field data to be used for the calibration and validation. The available aerial and
satellite imagery are commonly used for the mapping of seagrass landscapes and their application is
proven by various research papers. The imagery include satellite multi-spectral imagery (Landsat-TM,
ETM+) and aerial imagery from the Google Earth.
The overview of the collected data enables to summarize their following types:
\begin{itemize}
	\item[] 1. Optical spectra of P.oceanica, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig. \ref{fig:23})
	\item[]2. Aerial imagery from the \href{http://www.google.com/earth/index.html}{Google Earth}
	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:68})
	\item[]4. Results of underwater videometric measurements of the Olympus cameras made during the ship route
\end{itemize}
The available satellite imagery (Fig.\ref{fig:68}) and aerial images are read into the ArcGIS project.
The available broadband and hyperspectral remote sensing data are used for the mapping of the
seagrass in shelf areas no deeper than 4.0 meters, for the environmental monitoring in order to detect
the dynamics in changes of the seagrass distribution along Crete during the past 10 years using
different satellite images for 2000-2010.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.40]{UML_Dia.jpg}
	\caption{Workflow for the data aquisition, \href{http://live.gnome.org/Dia}{Dia} display}
	\label{fig:21}
\end{figure}

The spatial resolution of Landsat ETM+ image is 30 m in the visible and near infrared bands (bands
1-5 and 7); the spatial resolution of ASTER 15 m for the visible and near-infrared bands. IKONOS
acquires data in 3 visible channels and NIR, with spatial resolution of 1-4 meters. ASTER and
IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
research includes fieldwork in-situ measurements of the seagrass distribution along the northern and
south-eastern coasts of Crete in chosen locations.\\
The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
the seafloor have been received by the means of the optical sensors Trios-RAMSES Hyperspectral
UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig. \ref{fig:10}), both adjusted for the measurements of the irradiance and radiance (see Appendices: Tab. \ref{tab:18}, Tab. \ref{tab:19}).
The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:49}) and 430-650 for the water radiance (Fig.\ref{fig:47}). 
Afterwards, the measured values of the
radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
properties of the sea water and bottom cover types (Fig.\ref{fig:23}). The spectral range of radiance
cover diapason of 320-950 nm, and ittadiance measurements are covered in the interval of 280-500
nm, which is suitable for a characteristics of seagrass reflectance.
Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:43}, or \ref{fig:45}, enlarged) represent several series of
the measurements. The values of the spectral reflectance are received from the
computations of these values using Excel mathematical formulae. The graphs shown on Fig.\ref{fig:47}, Fig. \ref{fig:48}, Fig. \ref{fig:49} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively. Graph on Fig \ref{fig:44} displays statistuical analysis of the measured sets of observations, i.e. shows the midspread of the statistical quartiles (Q1-Q3), mean and extreme values.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{GNU-17.jpg}
	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:23}
\end{figure}
\pagebreak

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-18.jpg}
	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentray graphs below show the results of the statistical analysis.}
	\label{fig:26}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-21.jpg}
	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentray graphs below show the results of the statistical analysis}
	\label{fig:27}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-20.jpg}
	\caption{Statitical comparison of spectral reflectances of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
	\label{fig:28}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.40]{GNU-19.jpg}
	\caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and 			measured values (dotted): multiplot of measurement sets 200-300. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:33}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.45]{GNU-13.jpg}
		\caption{Multiplot of spectral reflectance of P.oceanica. Series 1-100. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
		\label{fig:35}
	\end{center}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[scale=0.5]{GNU-16.jpg}
		\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
		\label{fig:36}
	\end{center}
\end{figure}

@@

\begin{figure}
	\centering
	\subfloat [small-scaled (1:54,000)]{\label{fig:35a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
	\subfloat [enlarged (1:10,000)]{\label{fig:35b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
	\caption{Seafloor types of the Arina beach: \textit{P.oceanica,} sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
	\label{fig:37}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.40]{Fig-36.jpg}
		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }
		\label{fig:38}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.25]{Fig-37.jpg}
		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout}
		\label{fig:39}
	\end{center}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfloat [Meadows of the seagrass. Source: \href{http://www.google.com/earth/index.html}{Google Earth}]{\label{fig:38a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
	\subfloat [Patches (mattes) of the seagrass. Source: in-situ videometric measurements]{\label{fig:38b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
	\caption{Variations in spatial structure of the seagrass landscapes}
	\label{fig:40}
\end{figure}

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.25]{Fig-39.jpg}
		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project}
		\label{fig:41}
	\end{center}
\end{figure}

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.20]{App-1.jpg}
		\caption{Values of spectral reflectances of seagrass P.oceanica (fragment)}
		\label{fig:42}
	\end{center}
\end{figure}
\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.20]{GNU-401-420.jpg}
		\caption{Spectral reflectance of P.oceanica. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
		\label{fig:43}
	\end{center}
\end{figure}
\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.20]{GNU-401-420-candles.jpg}
		\caption{Statistical analysis of the measurement data: spectral reflectance of the P.oceanica. 
		Visualisation of the interquartile ranging. Example of data set 401-420. Shown midspread of  statistical quartiles Q1 and Q3, min and max values within the range. \href{http://www.gnuplot.info/}{Gnuplot}}
		\label{fig:44}
	\end{center}
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.20]{GNU-enlarged-M-401-420.jpg}
		\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
		Visualisation of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
		\label{fig:45}
	\end{center}
\end{figure}
\pagebreak
%%%

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral 			reflectance of \textbf{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline\hline
			\textbf{R} & \textbf{R square} & \textbf{Adjusted R square} & \textbf{Std. Error of the Estimate} \\ \hline\hline
			.277 1 & .077 & .068126692 & .462 \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:5}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
	\begin{center}
		\begin{tabular}{|c|c|c|c| c|c|}
			\hline\hline
			& \textbf{Sum of squares} & \textbf{df} & \textbf{Mean Square} & F & Sig.\\ \hline\hline
			Regression & 1.810 & 1 & 1.810 & 8.477 & .004\\ \hline
			Residual & 21.781 & 102 & .214 & & \\ \hline
			Total & 23.592 & 103 & & & \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:6}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
	\begin{center}
		\begin{tabular}{| c | c | p{2cm}| c | c | c |}
			\hline\hline
			& \multicolumn{2}{|c|}{Unstandardized Coefficients}\\
			\cline{2-3}
			& B & Std.Error & Stand. Coef. Beta & t & Sig.\\ \hline\hline
			wavelength & -.001 & .000 & -.277 & -2.912 & .004 \\ \hline
			constant & .464 & .121 & & 3.821 & .000 \\ \hline
		\end{tabular}
	\end{center}
	The underlying process assumed is independence (white noise).\\
	 Based on the asymptotic chi-square approximation.
	\label{tab:7}
\end{table}
\pagebreak

\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=0.20]{GNU-Stat-M-401-420.jpg}
	\caption{Fragment of the statistical analysis of the P.oceanica reflectance. Example of data set 401-420. 
		Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:46}
	\end{center}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.25]{GNU-Radiance_Bezier.jpg}
	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:47}
\end{figure}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters. \href{http://www.spss.com/}{SPSS}}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline\hline
		\textbf{Source of Variation} & \textbf{SS} & \textbf{df} & \textbf{MS} & \textbf{F} & \textbf{P value} & \textbf{F crit} \\ \hline
		Between Groups & 373841.7048 & 2 & 186920.8524 & 407.85359 & 1.11677 & 3.01153 \\ \hline
		Within Groups & 261233.1668 & 570 & 458.3038014 & & & \\ \hline
		Total &  635074.8716 & 572 & & & & \\ \hline
	\end{tabular}
	\end{center} 
	P more than .05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
	\label{tab:8}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters. \href{http://www.spss.com/}{SPSS}}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|}
		\hline\hline
		\textbf{Groups} & \textbf{Count} & \textbf{Sum} & \textbf{Average} & \textbf{Variance} \\ \hline\hline
		Column 1 & 191 &	760.126692 & 3.9797209 & 18.6948046 \\ \hline
		Column 2 & 191 &	4527.214218 & 23.7026922  & 251.4863978 \\ \hline
		Column 3 & 191 &	12465.19972 & 65.2628257 & 1104.7302015 \\ \hline
	\end{tabular}
	\end{center}
	P$>$0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
 	\label{tab:9}
\end{table}
\pagebreak

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-Irr-Splines.jpg}
		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}­}
		\label{fig:48}
	\end{center}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.25]{GNU-Irr.jpg}
	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:49}
\end{figure}
\pagebreak
\newpage
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-gamma-rad.jpg}
		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}}
		\label{fig:50}
	\end{center}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{G-normal-dist.jpg}
	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}­}
	\label{fig:51}
\end{figure}
